---
title: "American Woodcock"
author: "Amanda Klisavage"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, cache=TRUE)
```

## American Woodcock, *Scolopax minor*


## Methods

```{r libraries}
library(raster)
library(spocc)
library(RColorBrewer)
library(leaflet)
library(ggplot2)
library(tidyverse)
library(knitr)
library(sp)
library(sf)
library(plotly)
library(leaflet.extras)
library(htmlwidgets)
```


This project pulls species distribution data from R package [spocc](https://cran.r-project.org/web/packages/spocc/spocc.pdf) and  extracts [world climate data](http://worldclim.org/version2) via [Raster](https://cran.r-project.org/web/packages/raster/index.html)::[getData](https://www.rdocumentation.org/packages/raster/versions/3.0-7/topics/getData). Interactive maps are created in  [Leaflet](https://rstudio.github.io/leaflet/).

### Downloading Woodcock Data
spocc::occ is used to fetch occurrence data from Global Biodiversity Information Facility, or [GBIF](https://www.gbif.org/).

spocc::occ2df() converts the occurrence data into a data frame containing latitude, longitude, date, and a unique key. dplyr::mutate() is used to add a common name column.


```{r occ}
WCocc<-occ(query="Scolopax minor", limit=1000, has_coords=TRUE)
WCdf<-occ2df(WCocc)%>%
  mutate(common.name="American Woodcock")
```


### Downloading Climate Data

```{r climate}
precip <- getData(name = "worldclim", var="prec", res=5)
gain(precip)=0.1
prec_annual<-max(precip)
names(prec_annual) <- "pmax"

tmax_monthly <- getData(name = "worldclim", var="tmax", res=5)
gain(tmax_monthly)=0.1
tmax_annual<-max(tmax_monthly)
names(tmax_annual) <- "tmax"


e1<-extent(c(-135.001953,-45.966176,20.478516,55))

prec_US<-crop(prec_annual, e1)
temp_US<-crop(tmax_annual, e1)
```

### Preparing Occurence Data for Climate Data Extraction

[sp::SpatialPoints](https://www.rdocumentation.org/packages/sp/versions/1.3-1/topics/SpatialPoints) is first used to create an sp object from occurrence data for each species. This is necessary in order to extract climate data for each occurrence. [sf::st_as_sf](https://www.rdocumentation.org/packages/sf/versions/0.8-0/topics/st_as_sf) then converts the occurrence/climate data to an sf object. They are then spatially joined via [sf::st_join](https://www.rdocumentation.org/packages/sf/versions/0.8-0/topics/st_join).

```{r sp}
WCsp<-SpatialPoints(cbind(WCdf$longitude, WCdf$latitude))
WCprec<-raster::extract(prec_annual, WCsp, fun=max, na.rm=TRUE, small=TRUE, sp=TRUE)%>%
  st_as_sf()

WCtmax<-raster::extract(tmax_annual, WCsp, fun=max, na.rm=TRUE, small=TRUE, sp=TRUE)%>%
  st_as_sf()

WCUSclim<-st_join(WCprec, WCtmax)
```


## Results

### Interactive Leaflet Map



```{r leaflet}
pal <- colorNumeric(c("#FFFFCC", "#41B6C4", "#0C2C84"), values(prec_US),
  na.color = "transparent")
tpal<-colorNumeric(c("#0000FF","#00FF00","#FF0000"), values(temp_US), na.color="transparent")

map1 <- leaflet()%>%
  addProviderTiles("Esri.WorldPhysical")%>%
  addCircleMarkers(lng=WCdf$longitude, lat=WCdf$latitude, fillOpacity = 0.2, radius=2)%>%
  addRasterImage(prec_US, colors=pal, group="Precip")%>%
  addRasterImage(temp_US, colors="Spectral", group="Max Temp")%>%
  addLegend(pal = pal, values = values(prec_US),
    title = "Annual precipitation (mm)")%>%
 addLegend(pal = tpal, values = values(temp_US),
    title = "Maximum Temperature (C°)")%>%
  addLayersControl(overlayGroups = c("Precip", "Max Temp"),
    options = layersControlOptions(collapsed = FALSE))
map1
```

```


### Environmental Space Plots

```{r plot}
g<-ggplot(data=WCUSclim, aes(x=tmax, y=pmax, alpha=0.2))+
  geom_point()+
  ggtitle("Range of the American Woodcock")+
  ylab("Annual Precipitation, mm")+
  xlab("Maximum Annual Temperature, °C")+
  theme(legend.position = "none")


g
```



